Exploring Design Space for Differentially Private Bayesian Inference

### 2019-01-05 ICML2019/main.tex
update the section 3

The beta-binomial model can be immediately generalized to the Dirichlet-multinomial model, with underlying data multinomially distributed.
The \emph{bias} is represented by the parameter $\vtheta$, the vector of parameters of a categorically distributed random variable. The prior distribution over $\vtheta\in [0,1]^{k}$
is given by a Dirichelet distribution, $\dirichlet{n}{\valpha}$, for $k\in\mathbb{N}$,
and $\valpha\in(\mathbb{R}^{+})^{k}$, with p.d.f:
\[
  \Pr(\vtheta)\equiv\frac{1}{\mbetaf(\valpha)}\cdot \displaystyle\prod_{i=1}^{k}{\theta_i^{\alpha_i-1}}
\]
where $\mbetaf(\cdot)$ is the generalized beta function.
The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ values
coming from a universe $\datauni$, such that $\mid\datauni \mid=k$.
The likelihood function will be:
\[
  \Pr(\dataobs|\vtheta)\equiv \displaystyle\prod_{a_i\in\datauni}\theta_{i}^{\Delta \alpha_i},
\]
with $\Delta \alpha_i=\displaystyle\sum_{j=1}^{n}\iverson{x_j=a_i}$, where $\iverson{\cdot}$ represents Iverson bracket notation.
Denoting by $\Delta\valpha$ the vector $(\Delta\alpha_1,\dots \Delta\alpha_k)$ the posterior distribution over $\vtheta$ turns out to be
\[
   \bysinfer(\multinomial{n}{\vtheta}, \dirichlet{\valpha}, \dataobs)=\dirichlet{\valpha + \Delta \valpha} 
\]
where $+$ denotes the componentwise sum of vectors of reals.

@@

\section{Mechanism Proposition}
\label{sec_mechs}

\subsection{Laplace Mechanism Family}

\subsubsection{using $\ell_1$ norm metric}
 Adding noise to the posterior distribution parameters directly, through Laplace mechanism with post-processing, $\lapmech(\bysinfer, \dataobs)$, producing an output:
$$
\begin{array}{rcl}
& \lapmech(\bysinfer, \dataobs) \\
& = \betad(\alpha +  \lfloor{\Delta \alpha + Y}\rfloor^n_0,\beta + n - \lfloor{\Delta \alpha + Y}\rfloor^n_0),
\end{array} 
$$
 where $Y \thicksim \lap{0}{\frac{\Delta \bysinfer}{\epsilon}}$, a Laplace distribution with location $0$ and scale $\frac{\Delta \bysinfer}{\epsilon}$ in Beta-binomial model; and
$$
\begin{array}{rcl}
 & \lapmech(\bysinfer, \dataobs) = \\
 & \dirichlet(\alpha_1 +  \lfloor{\Delta \alpha_1 + Y_1}\rfloor^n_0, \cdots, \\
 & \alpha_k + \lfloor n - \sum_{i = 1}^{k-1}\lfloor{\Delta \alpha_i + Y_i}\rfloor^n_0 \rfloor^n_0),
\end{array} 
$$
 where $Y_i \thicksim \lap{0}{\frac{\Delta \bysinfer}{\epsilon}}$ in Dirichlet-multinomial model. $\lfloor \cdot \rfloor^n_0$ is taking the floor value and truncating into $[0,n]$ to make sure the noised posterior is valid.

 We use $ \lapmech(\dataobs) = \betad(\tilde{\alpha}, \tilde{\beta})$ or $\dirichlet(\tilde{\valpha})$ for short. Then release it as the private posterior distribution.

 The sensitivity of the inference process w.r.t. $\ell_1$ norm in $\lapmech$ is:
 \[
 \Delta \bysinfer \equiv \max\limits_{\dataobs, \dataobs' \in \{0,1\}^{n}, ||\dataobs - \dataobs'||_1 \leq 1} ||\bysinfer(\dataobs) - \bysinfer(\dataobs')||_1,
 \]
 which is proportional to the dimensionality.

\subsubsection{using improved $\ell_1$ norm metric}
 Noise added to posterior distribution parameters are scaled to smaller sensitivity in this improved Laplace mechanism. Because in terms of two adjacent data sets $\dataobs, \dataobs'$, their posterior distributions by Bayesian inference -- $\bysinfer(\dataobs), \bysinfer(\dataobs')$ -- which parameter differs at most in 2 dimensions even though extended to Dirichlet-multinomial mode, i.e., $\Delta \bysinfer \leq 2$.

 Then it is enough to use sensitivity $1$ in 2 dimensions and $2$ in higher dimensions, then the improved Laplace mechanism $\ilapmech(\bysinfer, \dataobs)$ is producing an output:

$$
\begin{array}{rcl}
& \ilapmech(\bysinfer, \dataobs) \\
& = \betad(\alpha +  \lfloor{\Delta \alpha + Y}\rfloor^n_0,\beta + n - \lfloor{\Delta \alpha + Y}\rfloor^n_0),
\end{array} 
$$
 where $Y \thicksim \lap{0}{\frac{1}{\epsilon}}$ in Beta-binomial model; and
$$
\begin{array}{rcl}
 & \ilapmech(\bysinfer, \dataobs) = \\
 & \dirichlet(\alpha_1 +  \lfloor{\Delta \alpha_1 + Y_1}\rfloor^n_0, \cdots, \\
 & \alpha_k + \lfloor n - \sum_{i = 1}^{k-1}\lfloor{\Delta \alpha_i + Y_i}\rfloor^n_0 \rfloor^n_0),
\end{array} 
$$
where $Y_i \thicksim \lap{0}{\frac{2}{\epsilon}}$ in Dirichlet-multinomial model.

Both Laplace mechanism and improved one are $\epsilon -$differential privacy\cite{dwork2014algorithmic}.

>>>>

The beta-binomial model can be immediately generalized to the Dirichlet-multinomial model, with underlying data multinomially distributed.
The \emph{bias} is represented by the parameter $\vtheta$, the vector of parameters of a categorically distributed random variable. The prior distribution over $\vtheta\in [0,1]^{k}$
is given by a Dirichlet distribution, $\dirichlet{\valpha}$, for $k\in\mathbb{N}$,
and $\valpha\in(\mathbb{R}^{+})^{k}$, with p.d.f:
\[
  \Pr(\vtheta)\equiv\frac{1}{\mbetaf(\valpha)}\cdot \displaystyle\prod_{i=1}^{k}{\theta_i^{\alpha_i-1}}
\]
where $\mbetaf(\cdot)$ is the generalized beta function.
The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ values
coming from a universe $\datauni$, such that $\mid\datauni \mid=k$.
The likelihood function will be:
\[
  \Pr(\dataobs|\vtheta)\equiv \displaystyle\prod_{a_i\in\datauni}\theta_{i}^{\Delta \alpha_i},
\]
with $\Delta \alpha_i=\displaystyle\sum_{j=1}^{n}\iverson{x_j=a_i}$, where $\iverson{\cdot}$ represents Iverson bracket notation.
Denoting by $\Delta\valpha$ the vector $(\Delta\alpha_1,\dots \Delta\alpha_k)$ the posterior distribution over $\vtheta$ turns out to be
\[
   \bysinfer(\multinomial{n}{\vtheta}, \dirichlet{\valpha}, \dataobs)=\dirichlet{\valpha + \Delta \valpha} 
\]
where $+$ denotes the componentwise sum of vectors of reals.

@@

\section{Mechanism Proposition}
\label{sec_mechs}
\subsection{Laplace Mechanism Family}

\subsubsection{Using $\ell_1$ Norm Metric}
  We firstly present the fully explored mechanism in Laplace family -- Laplace mechanism. This mechanism ($\lapmech(\dataobs,\bysinfer, \mathsf{prior}, \mathsf{likelihood}, \epsilon)$) simply adds Laplace noise with post-processing to the posterior distribution parameters, producing an output: $\betad(\alpha +  \lfloor{\Delta \alpha + \eta}\rfloor^n_0,\beta + n - \lfloor{\Delta \alpha + \eta}\rfloor^n_0)$ with $\eta \thicksim \lap{0}{\frac{\Delta \bysinfer}{\epsilon}}$, where $\lfloor \cdot \rfloor^n_0$ is taking the floor value and truncating into $[0,n]$ to make sure the noised posterior is valid, as shown in Mechanism \ref{mech:lapmech}.\\
  Since we are adding noise directly to the parameter, the sensitivity should be calibrated to $\ell_1$ of two Beta distributions:
  $
   \Delta \bysinfer \equiv \max\limits_{\dataobs, \dataobs' \in \{0,1\}^{n}, ||\dataobs - \dataobs'||_1 \leq 1} ||\bysinfer(\dataobs) - \bysinfer(\dataobs')||_1
  $,
  i.e., proportional to the dimensionality.

  \begin{algorithm}
  \caption{$\lapmech$ in Beta-binomial model}
  \label{mech:lapmech}
  \begin{algorithmic}
  \REQUIRE observed data set $\dataobs \in \{0,1\}^n$, inference algorithm: $\bysinfer$, 
  prior: $\betad(\alpha, \beta)$, likelihood: $\binomial{n}{\theta}$, $\epsilon$
  \STATE \quad {\bf let} $\betad(\alpha', \beta') = \bysinfer(\binomial{n}{\theta}, \betad(\alpha, \beta), \dataobs)$.  
  \STATE \quad {\bf let} $\eta \sim \lap{0}{\frac{2}{\epsilon}}$.
  \STATE \quad {\bf let} $\tilde{\alpha}=\alpha + \lfloor{(\alpha' - \alpha) + \eta}\rfloor^n_0$ 
  \STATE \quad {\bf let} $\tilde{\beta}=\beta + n - \lfloor{(\alpha' - \alpha) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \ENSURE $\betad(\tilde{\alpha}, \tilde{\beta})$
  \end{algorithmic}
  \end{algorithm}

  In Dirichlet-multinomial model, $\lapmech$ produces output: $ \dirichlet{\alpha_1 +  \lfloor{\Delta \alpha_1 + \eta_1}\rfloor^n_0, \cdots, \alpha_k + \lfloor n - \sum_{i = 1}^{k-1}\lfloor{\Delta \alpha_i + \eta_i}\rfloor^n_0 \rfloor^n_0},$
  where $\eta_i \thicksim \lap{0}{\frac{k - 1}{\epsilon}}$.
  We will use $ \lapmech(\dataobs) = \betad(\tilde{\alpha}, \tilde{\beta})$ or $\dirichlet{\tilde{\valpha}}$ for short.

\subsubsection{Using $\ell_1$ Norm with Improved Sensitivity}
 The $\lapmech$ presented above can be improved in our Bayesian inference model, by adding noise scaled to smaller sensitivity in this improved Laplace mechanism $\ilapmech$.\\
 In terms of two adjacent data sets $\dataobs, \dataobs'$, their posterior distributions by Bayesian inference -- $\bysinfer(\dataobs), \bysinfer(\dataobs')$ -- which parameter differs at most in 2 dimensions even though extended to Dirichlet-multinomial mode, i.e., $\Delta \bysinfer \leq 2$. As a results, it is enough to use sensitivity $1$ in 2 dimensions and $2$ in higher dimensions.\\
 By calibrating noise to this improved sensitivity, $\ilapmech(\dataobs,\bysinfer, \mathsf{prior}, \mathsf{likelihood}, \epsilon)$ produces output:
 $\betad(\alpha +  \lfloor{\Delta \alpha + \eta}\rfloor^n_0,\beta + n - \lfloor{\Delta \alpha + \eta}\rfloor^n_0)
$
with $\eta \thicksim \lap{0}{\frac{1}{\epsilon}}$ in Beta-binomial model and
$\dirichlet{\alpha_1 +  \lfloor{\Delta \alpha_1 + \eta_1}\rfloor^n_0, \cdots, 
\alpha_k + \lfloor n - \sum_{i = 1}^{k-1}\lfloor{\Delta \alpha_i + \eta_i}\rfloor^n_0 \rfloor^n_0}
$
with $\eta_i \thicksim \lap{0}{\frac{2}{\epsilon}}$ in Dirichlet-multinomial model. Details are shown in Mechanism \ref{mech:ilapmech}.
  \begin{algorithm}
  \caption{$\ilapmech$ in Dirichlet-multinomial model}
  \label{mech:ilapmech}
  \begin{algorithmic}
  \REQUIRE observed data set $\dataobs \in \mathcal{X}^n$, inference algorithm: $\bysinfer$, 
  prior: $\dirichlet{\valpha}$, likelihood: $\multinomial{n}{\vtheta}$, $\epsilon$
  \STATE \quad {\bf let} $\dirichlet{\valpha'} = \bysinfer(\multinomial{n}{\vtheta}, \dirichlet{\valpha}, \dataobs)$.
  \STATE \quad {\bf Initialize} $\tilde{\valpha} = (0,\dots )$, $k = |\mathcal{X}|$.
  \STATE \quad {\bf For} $i = 1 \dots (k - 1)$:
  \STATE \quad \quad  {\bf let} $\eta \sim \lap{0}{\frac{2}{\epsilon}}$.
  \STATE \quad \quad  $\tilde{\alpha_i}=\alpha_i + \lfloor{(\alpha_i' - \alpha_i) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \STATE \quad $\tilde{\alpha_k} = \alpha_k + \lfloor n - \sum_{i = 1}^{k-1}\lfloor{(\alpha_i' - \alpha_i) + \eta_i}\rfloor^n_0 \rfloor^n_0$
  \ENSURE $\dirichlet{\tilde{\valpha}}$
  \end{algorithmic}
  \end{algorithm}
